---
title: "Nass Data Summary"
---

```{r}
#| include: false

library(tidyverse)
library(reshape2)
```

## 1. Background

introduction to Nass watershed and salmon

## 2. Summary

example - modified some code from Joe's Shiny app to make a plot of harvests and escapement by year and CU - this is not very useful to look at, better if broken down by species and smaller timeframes

```{r}
#| warning: false
#| message: false

salmon <- readRDS("salmon.rds")
exploitation <- readRDS("exploitation.rds")

# Exploitation plot    
ggplot(exploitation, aes(x = Year, y = Fish/100000, fill = Proportion)) + 
  geom_col(position = "stack") +
  scale_fill_manual(values = c("gray78", "gray46")) +
  labs(x = "Year", y = "x100,000 Salmon", title = "Area 3 Exploitation by Year")+
  facet_wrap(~`Species Name`) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

# Salmon data plot

ggplot(salmon, aes(x = `CU Name`, y = Fish/100000, fill = Proportion)) +
  geom_col() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "x100,000 Salmon", x = "Conservation Unit", title = "Breakdown by CU")

```

}

## 3. Escapement Goals
simple plot of escapement goal and min (minimum?) goal for each species
```{r}
#| message: false
#| echo: false

esc_goals <- read_csv("data/nass salmon escapement goals.csv", show_col_types = FALSE)

esc_reshaped <- melt(esc_goals, id.vars = 'species')

ggplot(esc_reshaped, aes(x = species, y = value, fill=variable)) +
  geom_col(position='dodge') +
  ylab("Goal number") +
  xlab("Species")

```

## 4. Sockeye Salmon

a.  Area 3

b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## 5. Chinook Salmon

a.  Area 3

b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## Coho Salmon

a.  Area 3

b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## Pink Salmon

a.  Area 3

b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## Chum Salmon

a.  Area 3

b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## Steelhead Trout

a.  Area 3

b.  Conservation Units

c.  Streams

d.  Exploitation Rates
