---
title: "Nass Data Summary"
subtitle: "September 19, 2024"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---


```{r}
#| include: false

library(tidyverse)
library(reshape2)
library(ggpubr)
library(readxl)
library(data.table)
```

## 1. Background

introduction to Nass watershed and salmon

## 2. Summary

```{r}
#| warning: false
#| message: false
#| echo: false

salmon <- readRDS("salmon.rds")
exploitation <- readRDS("exploitation.rds")

# Exploitation plot    
ggplot(exploitation, aes(x = Year, y = Fish/100000, fill = Proportion)) + 
  geom_col(position = "stack") +
  scale_fill_manual(values = c("gray78", "gray46")) +
  labs(x = "Year", y = "x100,000 Salmon", title = "Area 3 Exploitation by Year")+
  facet_wrap(~`Species Name`) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

# Salmon data plot

ggplot(salmon, aes(x = `CU Name`, y = Fish/100000, fill = Proportion)) +
  geom_col() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "x100,000 Salmon", x = "Conservation Unit", title = "Breakdown by CU")

```

## 3. Escapement Goals

simple plot of escapement goal and minimum goal for each species

```{r}
#| message: false
#| echo: false
#| warning: false

esc_goals <- read_csv("data/nass salmon escapement goals.csv", show_col_types = FALSE)

esc_reshaped <- melt(esc_goals, id.vars = 'species')

ggplot(esc_reshaped, aes(x = species, y = value, fill=variable)) +
  geom_col(position='dodge') +
  ylab("Goal number") +
  xlab("Species")

```

## 4. Sockeye Salmon

a.  Area 3
```{r load and format data}
#| include: false
#| message: false
#| echo: false
#| warning: false

nass_mez <- fread("data/nass vs mez TR TE new.csv")
nass <- read_excel("data/nass data 240313.xlsx",sheet="meziadin and nass te and tr")
allyears<-fread("data/nass combined stat area data 240313.csv")

esc.plots <- allyears %>%
  filter(`Species Name`=="Sockeye")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

```

```{r sockeye esc plot}
#| warning: false
#| message: false
#| echo: false

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey40","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1980,2025))+
  theme_bw()+
  labs(y="Number of Sockeye Salmon")

er.plots<-allyears %>%
  filter(`Species Name`== "Sockeye",Year<=2023)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1980,2025)

ggarrange(pe,per,ncol=1,align="v")

```

b.  Conservation Units
```{r load CU data}
#| include: false
#| message: false
#| warning: false

CUs <- fread("data/NUSEDS escapement by CU 20220902.csv") %>%
  pivot_longer(`2021`:`1950`, names_to = "Year", values_to = "Escapement") %>%
  mutate(Year = as.numeric(Year))

(pe <- CUs %>%
  filter(SPECIES=="Sockeye")%>%
  ggplot(aes(x = CU_INDEX, y = Escapement))+
  geom_col()+
  #scale_fill_manual(values=c("grey90","grey40","blue"))+
  #guides(fill=guide_legend(title=""))+
  #xlim(c(1980,2025))+
  theme_bw()+
  labs(y="Number of Sockeye Salmon", x = "Conservation Unit"))
 # facet_wrap(~CU_NAME)

# colour by CU name?

```

c.  Streams

d.  Exploitation Rates

## 5. Chinook Salmon
a.  Area 3
```{r chinook esc plot}
#| message: false
#| echo: false
#| warning: false

esc.plots <- allyears %>%
  filter(`Species Name`== "Chinook")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey40","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1980,2025))+
  theme_bw()+
  labs(y="Number of Chinook Salmon")

er.plots<-allyears %>%
  filter(`Species Name`== "Chinook",Year<2024)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1980,2025)

ggarrange(pe,per,ncol=1,align="v")

```

b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## Coho Salmon
a.  Area 3
```{r coho esc plot}
#| message: false
#| echo: false
#| warning: false

esc.plots <- allyears %>%
  filter(`Species Name`== "Coho")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey40","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1980,2025))+
  theme_bw()+
  labs(y="Number of Coho Salmon")

er.plots<-allyears %>%
  filter(`Species Name`== "Coho",Year<2024)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1980,2025)

ggarrange(pe,per,ncol=1,align="v")


```
b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## Pink Salmon (even years)
a.  Area 3
```{r pink even esc plot}
#| message: false
#| echo: false
#| warning: false

esc.plots <- allyears %>%
  filter(`Species Name`=="Pink (even years)")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey20","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1960,2025))+
  theme_bw()+
  labs(y="Number of Pink Salmon")

er.plots<-allyears %>%
  filter(`Species Name`== "Pink (even years)",Year<2024)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1960,2025)

ggarrange(pe,per,ncol=1,align="v")

```

b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## Chum Salmon
a.  Area 3
```{r chum esc plot}
#| message: false
#| echo: false
#| warning: false

esc.plots <- allyears %>%
  filter(`Species Name`=="Chum")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey40","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1950,2025))+
  theme_bw()+
  labs(y="Number of Chum Salmon")

er.plots<-allyears %>%
  filter(`Species Name`=="Chum",Year<2024)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1950,2025)

ggarrange(pe,per,ncol=1,align="v")

```

b.  Conservation Units

c.  Streams

d.  Exploitation Rates

## Steelhead Trout
a.  Area 3
```{r load steelhead data}
#| include: false

nisgaash <- read_csv("data/nisgaa-tech-table-data.csv")%>%
  filter(Species == "Steelhead")%>%
  select(Year,TRTC, TR)%>%
  mutate(`SEAK Harvest`=TR-TRTC)%>%
  pivot_longer(c(`TRTC`,`SEAK Harvest`),names_to="type",values_to="fish")

```

```{r steelhead esc plot}
#| message: false
#| echo: false
#| warning: false

sh<-ggplot(nisgaash, aes(x = Year, y = fish,fill=ifelse(Year==2023,"Forecasted\nTRTC",type))) +
  geom_col(color="black")+
  scale_fill_manual(values=c("blue","grey90","grey40"))+
  guides(fill=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Number of Steelhead")+
  xlim(1990,2025)

sh

```

b.  Conservation Units

c.  Streams

d.  Exploitation Rates
