---
title: "Nass Data Summary"
subtitle: "September 19, 2024"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---


```{r}
#| include: false

library(tidyverse)
library(reshape2)
library(ggpubr)
library(readxl)
library(data.table)
library(gridExtra)
library(RColorBrewer)
```

## 1. Background

introduction to Nass watershed and salmon

## 2. Summary

```{r}
#| warning: false
#| message: false
#| echo: false

salmon <- readRDS("salmon.rds")
exploitation <- readRDS("exploitation.rds")

# Exploitation plot    
ggplot(exploitation, aes(x = Year, y = Fish/100000, fill = Proportion)) + 
  geom_col(position = "stack") +
  #scale_fill_manual(values = c("gray78", "gray46")) +
  labs(x = "Year", y = "x100,000 Salmon", title = "Area 3 Exploitation by Year")+
  facet_wrap(~`Species Name`) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))

# Salmon data plot

#ggplot(salmon, aes(x = `CU Name`, y = Fish/100000, fill = Proportion)) +
#  geom_col() +
#  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#  labs(y = "x100,000 Salmon", x = "Conservation Unit", title = "Breakdown by CU")

```

## 3. Escapement Goals

simple plot of escapement goal and minimum goal for each species

```{r}
#| message: false
#| echo: false
#| warning: false

esc_goals <- read_csv("data/nass salmon escapement goals.csv", show_col_types = FALSE)

esc_reshaped <- melt(esc_goals, id.vars = 'species')

ggplot(esc_reshaped, aes(x = species, y = value, fill=variable)) +
  geom_col(position='dodge') +
  ylab("Goal number") +
  xlab("Species")

```

## 4. Sockeye Salmon

a.  Area 3
```{r load and format data}
#| include: false
#| message: false
#| echo: false
#| warning: false

nass_mez <- fread("data/nass vs mez TR TE new.csv")
nass <- read_excel("data/nass data 240313.xlsx",sheet="meziadin and nass te and tr")
allyears<-fread("data/nass combined stat area data 240313.csv")

esc.plots <- allyears %>%
  filter(`Species Name`=="Sockeye")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

```

```{r sockeye esc plot}
#| warning: false
#| message: false
#| echo: false

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey40","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1980,2025))+
  theme_bw()+
  labs(y="Number of Sockeye Salmon")

er.plots<-allyears %>%
  filter(`Species Name`== "Sockeye",Year<=2023)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1980,2025)

ggarrange(pe,per,ncol=1,align="v")

```

b.  Conservation Units
```{r load CU data}
#| include: false
#| message: false
#| warning: false

CUs <- fread("data/NUSEDS escapement by CU 20220902.csv") %>%
  pivot_longer(`2021`:`1950`, names_to = "Year", values_to = "Escapement") %>%
  mutate(Year = as.numeric(Year))


area3 <- CUs %>%
  filter(AREA == "3B" | AREA == "3A") %>%
  filter(SPECIES=="Sockeye")
```

```{r plot annual escapement by CU}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 2

# for (i in 1:length(unique(area3$CU_NAME))){
#   CU <- unique(area3$CU_NAME)[i]
#   filtered_data <- area3 %>% filter(CU_NAME == CU)
#   pl <- ggplot(filtered_data, aes(x = Year, y = Escapement))+
#     geom_col(fill = "grey60")+
#     xlim(c(1980,2025))+
#     theme_bw()+
#     labs(y="Number of Sockeye Salmon", x = "Year",
#          title = paste("Annual Sockeye Escapement for", CU, "Conservation Unit"))
#   print(pl)
# }

```

c.  Streams
```{r plot annual escapement by CU and stream}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 1

nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)

area3$WATERBODY  <- with(area3, reorder(WATERBODY, Escapement))

for (i in 1:length(unique(area3$CU_NAME))){
  CU <- unique(area3$CU_NAME)[i]
  filtered_data <- area3 %>% filter(CU_NAME == CU)
  pl <- ggplot(filtered_data, aes(x = Year, y = Escapement, fill = filtered_data$WATERBODY))+
    geom_col(color = "black")+
    scale_fill_manual(values = mycolors) +
    xlim(c(1980,2025))+
    theme_bw()+
    labs(y="Number of Sockeye Salmon", x = "Year",
         title = paste("Annual Sockeye Escapement for",CU, "CU, by Stream"), fill = "Stream name")
  print(pl)
}

```
d.  Exploitation Rates

## 5. Chinook Salmon
a.  Area 3
```{r chinook esc plot}
#| message: false
#| echo: false
#| warning: false

esc.plots <- allyears %>%
  filter(`Species Name`== "Chinook")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey40","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1980,2025))+
  theme_bw()+
  labs(y="Number of Chinook Salmon")

er.plots<-allyears %>%
  filter(`Species Name`== "Chinook",Year<2024)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1980,2025)

ggarrange(pe,per,ncol=1,align="v")

```

b.  Conservation Units
```{r load CU data chinook}
#| include: false
#| message: false
#| warning: false

CUs <- fread("data/NUSEDS escapement by CU 20220902.csv") %>%
  pivot_longer(`2021`:`1950`, names_to = "Year", values_to = "Escapement") %>%
  mutate(Year = as.numeric(Year))


area3 <- CUs %>%
  filter(AREA == "3B" | AREA == "3A") %>%
  filter(SPECIES=="Chinook")
```

```{r plot annual escapement by CU chinook}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 2

# for (i in 1:length(unique(area3$CU_NAME))){
#   CU <- unique(area3$CU_NAME)[i]
#   filtered_data <- area3 %>% filter(CU_NAME == CU)
#   pl <- ggplot(filtered_data, aes(x = Year, y = Escapement))+
#     geom_col(color = "black", fill = "grey90")+
#     xlim(c(1980,2025))+
#     theme_bw()+
#     labs(y="Number of Chinook", x = "Year",
#          title = paste("Annual Chinook Escapement for", CU, "Conservation Unit"))
#   print(pl)
# }

```
c.  Streams
```{r plot chinook annual escapement by CU and stream}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 1

nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)

for (i in 1:length(unique(area3$CU_NAME))){
  CU <- unique(area3$CU_NAME)[i]
  filtered_data <- area3 %>% filter(CU_NAME == CU)
  pl <- ggplot(filtered_data, aes(x = Year, y = Escapement, fill = filtered_data$WATERBODY))+
    geom_col(color = "black")+
    scale_fill_manual(values = mycolors) +
    xlim(c(1980,2025))+
    theme_bw()+
    labs(y="Number of Chinook Salmon", x = "Year",
         title = paste("Annual Chinook Escapement for",CU, "CU, by Stream"), fill = "Stream name")
  print(pl)
}

```
d.  Exploitation Rates

## Coho Salmon
a.  Area 3
```{r coho esc plot}
#| message: false
#| echo: false
#| warning: false

esc.plots <- allyears %>%
  filter(`Species Name`== "Coho")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey40","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1980,2025))+
  theme_bw()+
  labs(y="Number of Coho Salmon")

er.plots<-allyears %>%
  filter(`Species Name`== "Coho",Year<2024)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1980,2025)

ggarrange(pe,per,ncol=1,align="v")


```
b.  Conservation Units
```{r load CU data coho}
#| include: false
#| message: false
#| warning: false

CUs <- fread("data/NUSEDS escapement by CU 20220902.csv") %>%
  pivot_longer(`2021`:`1950`, names_to = "Year", values_to = "Escapement") %>%
  mutate(Year = as.numeric(Year))


area3 <- CUs %>%
  filter(AREA == "3B" | AREA == "3A") %>%
  filter(SPECIES=="Coho")
```

```{r plot annual escapement by CU coho}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 2

# for (i in 1:length(unique(area3$CU_NAME))){
#   CU <- unique(area3$CU_NAME)[i]
#   filtered_data <- area3 %>% filter(CU_NAME == CU)
#   pl <- ggplot(filtered_data, aes(x = Year, y = Escapement))+
#     geom_col(color = "black", fill = "grey90")+
#     xlim(c(1980,2025))+
#     theme_bw()+
#     labs(y="Number of Coho", x = "Year",
#          title = paste("Annual Coho Escapement for", CU, "Conservation Unit"))
#   print(pl)
# }

```
c.  Streams
```{r plot coho annual escapement by CU and stream}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 1


for (i in 1:length(unique(area3$CU_NAME))){
  CU <- unique(area3$CU_NAME)[i]
  filtered_data <- area3 %>% filter(CU_NAME == CU)
  pl <- ggplot(filtered_data, aes(x = Year, y = Escapement, fill = filtered_data$WATERBODY))+
    geom_col(color = "black")+
    scale_fill_brewer(palette = "Paired") +
    xlim(c(1980,2025))+
    theme_bw()+
    labs(y="Number of Coho Salmon", x = "Year",
         title = paste("Annual Coho Escapement for",CU, "CU, by Stream"), fill = "Stream name")
  print(pl)
}

```
d.  Exploitation Rates

## Pink Salmon (even years)
a.  Area 3
```{r pink even esc plot}
#| message: false
#| echo: false
#| warning: false

esc.plots <- allyears %>%
  filter(`Species Name`=="Pink (even years)")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey20","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1960,2025))+
  theme_bw()+
  labs(y="Number of Pink Salmon")

er.plots<-allyears %>%
  filter(`Species Name`== "Pink (even years)",Year<2024)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1960,2025)

ggarrange(pe,per,ncol=1,align="v")

```

b.  Conservation Units
```{r load CU data pink}
#| include: false
#| message: false
#| warning: false

CUs <- fread("data/NUSEDS escapement by CU 20220902.csv") %>%
  pivot_longer(`2021`:`1950`, names_to = "Year", values_to = "Escapement") %>%
  mutate(Year = as.numeric(Year))


area3 <- CUs %>%
  filter(AREA == "3B" | AREA == "3A") %>%
  filter(SPECIES=="Pink") %>%
  filter(!is.na(Escapement))
```

```{r plot annual escapement by CU pink}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 1

# for (i in 1:length(unique(area3$CU_NAME))){
#   CU <- unique(area3$CU_NAME)[i]
#   filtered_data <- area3 %>% filter(CU_NAME == CU)
#   pl <- ggplot(filtered_data, aes(x = Year, y = Escapement))+
#     geom_col(fill = "grey60")+
#     xlim(c(1980,2025))+
#     theme_bw()+
#     labs(y="Number of Pink Salmon", x = "Year",
#          title = paste("Annual Pink Salmon Escapement for", CU, "Conservation Unit"))
#   print(pl)
# }

```
c.  Streams
```{r plot pink annual escapement by CU and stream}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 1


for (i in 1:length(unique(area3$CU_NAME))){
  CU <- unique(area3$CU_NAME)[i]
  filtered_data <- area3 %>% filter(CU_NAME == CU)
  pl <- ggplot(filtered_data, aes(x = Year, y = Escapement, fill = filtered_data$WATERBODY))+
    geom_col(color = "black")+
    scale_fill_brewer(palette = "Paired") +
    xlim(c(1980,2025))+
    theme_bw()+
    labs(y="Number of Pink Salmon", x = "Year",
         title = paste("Annual Pink Escapement for",CU, "CU, by Stream"), fill = "Stream name")
  print(pl)
}

```
d.  Exploitation Rates

## Chum Salmon
a.  Area 3
```{r chum esc plot}
#| message: false
#| echo: false
#| warning: false

esc.plots <- allyears %>%
  filter(`Species Name`=="Chum")%>%
  pivot_longer(c(`Total Harvest`, `Escapement`), names_to = "Proportion", values_to = "Fish")%>%
  mutate(Type=ifelse(Year==2024,"Forecasted\nTRTC",Proportion))

pe<-ggplot(esc.plots, aes(x = Year, y = Fish, fill = factor(Type,levels=c("Total Harvest","Escapement","Forecasted\nTRTC"))))+
  geom_col(color="black",width = 1)+
  scale_fill_manual(values=c("grey90","grey40","blue"))+
  guides(fill=guide_legend(title=""))+
  xlim(c(1950,2025))+
  theme_bw()+
  labs(y="Number of Chum Salmon")

er.plots<-allyears %>%
  filter(`Species Name`=="Chum",Year<2024)%>%
  pivot_longer(c(`SEAK ER`, `CDN ER`,`Total ER`), names_to = "Area", values_to = "ER")

per<-ggplot(er.plots, aes(x = Year, y = ER,color=Area)) +
  geom_point()+geom_line()+
  scale_color_manual(values=c("red","blue","black"))+
  guides(color=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Exploitation Rate")+
  ylim(0,1)+
  xlim(1950,2025)

ggarrange(pe,per,ncol=1,align="v")

```

b.  Conservation Units
```{r load CU data chum}
#| include: false
#| message: false
#| warning: false

CUs <- fread("data/NUSEDS escapement by CU 20220902.csv") %>%
  pivot_longer(`2021`:`1950`, names_to = "Year", values_to = "Escapement") %>%
  mutate(Year = as.numeric(Year))


area3 <- CUs %>%
  filter(AREA == "3B" | AREA == "3A") %>%
  filter(SPECIES=="Chum")
```

```{r plot annual escapement by CU chum}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 2

# for (i in 1:length(unique(area3$CU_NAME))){
#   CU <- unique(area3$CU_NAME)[i]
#   filtered_data <- area3 %>% filter(CU_NAME == CU)
#   pl <- ggplot(filtered_data, aes(x = Year, y = Escapement))+
#     geom_col(fill = "grey60")+
#     xlim(c(1980,2025))+
#     theme_bw()+
#     labs(y="Number of Chum Salmon", x = "Year",
#          title = paste("Annual Chum Escapement for", CU, "Conservation Unit"))
#   print(pl)
# }

```
c.  Streams
```{r plot chum annual escapement by CU and stream}
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 1

nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)

for (i in 1:length(unique(area3$CU_NAME))){
  CU <- unique(area3$CU_NAME)[i]
  filtered_data <- area3 %>% filter(CU_NAME == CU)
  pl <- ggplot(filtered_data, aes(x = Year, y = Escapement, fill = filtered_data$WATERBODY))+
    geom_col(color = "black")+
    scale_fill_manual(values = mycolors) +
    xlim(c(1980,2025))+
    theme_bw()+
    labs(y="Number of Chum Salmon", x = "Year",
         title = paste("Annual Chum Escapement for",CU, "CU, by Stream"), fill = "Stream name")
  print(pl)
}

```
d.  Exploitation Rates

## Steelhead Trout
a.  Area 3
```{r load steelhead data}
#| include: false

nisgaash <- read_csv("data/nisgaa-tech-table-data.csv")%>%
  filter(Species == "Steelhead")%>%
  select(Year,TRTC, TR)%>%
  mutate(`SEAK Harvest`=TR-TRTC)%>%
  pivot_longer(c(`TRTC`,`SEAK Harvest`),names_to="type",values_to="fish")

```

```{r steelhead esc plot}
#| message: false
#| echo: false
#| warning: false

sh<-ggplot(nisgaash, aes(x = Year, y = fish,fill=ifelse(Year==2023,"Forecasted\nTRTC",type))) +
  geom_col(color="black")+
  scale_fill_manual(values=c("blue","grey90","grey40"))+
  guides(fill=guide_legend(title=NULL))+
  theme_bw()+
  labs(y="Number of Steelhead")+
  xlim(1990,2025)

sh

```

b.  Conservation Units

c.  Streams

d.  Exploitation Rates
